Making multiple-locus association mapping on a genome-wide scale routine

### 2017-11-09 AM+draft.tex
09Nov17 11am commit of methods under development

We compared the performance of Eagle against five other multiple locus association mapping methods: BigRR, 
glmnet, FaST-LASSO, MMLM, and r2VIM.  These five methods were selected because of their demonstrated value 
for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
features is given in Table X.

BigRR, FaST-LASSO, and glmnet are all based on regularization methods. Regression based regularization methods 
are well suited to solving problems where the number of predictors (p) is far greater than the number of samples (n).  It 
should come as no surprise then that multiple locus association mapping are dominated by regularization-based approaches.
Regularisation methods are able to yield solutions to problems that OLS cannot. They introduce a penalty function that balances
bias against reduced variance.  BigRR, an R package, implements gen

BigRR is an R package that implements

>>>>

We compared the computational and statistical performance of Eagle against five other multiple locus applications for 
association mapping  These were BigRR, LMM-Lasso, glmnet, MLMM  and, r2VIM. 
BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. Regularisation methods have dominated 
the development of multiple locus approaches.  They give opportunity for the effect of all the snp to be modelled simultaneously. 
By introducing a penalty function that balances bias and variance, regularisation methods can 
handle parameter estimation problems where the number of predictors is far greater than the number of samples.  
MLMM is similar in philosophy to Eagle in that it treats association mapping as a model building problem. Here, a linear mixed 
model is built through a forward-backward selected strategy.  
R2VIM is different to the others in that it implements 
random forests for association mapping. 
Random forests is a machine learning approach that forms an ensemble of decision trees for regression and 
classification. 
These five methods were selected because of their demonstrated value 
for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
features is given in Table X. Each method is now described in a little more detail.

BigRR is an R package that implements generalized ridge regression. Generalized ridge regression is 
 an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
a more realistic model to be specified. The model used in BigRR is 
$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.

BigRR is an R package that implements generalized ridge regression
which is an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
a more realistic model to be specified. The model used in BigRR is 
$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.

BigRR, LMM-Lasso, and glmnet all implement regularization methods for parameter
estimation. Regularization methods are well suited to problems where the number of predictors is far greater than the 
number of samples. It is not surprising then that regularization methods have attracted considerable interest, primarily for 
genomic prediction (refs) but also association mapping (refs). 
BigRR implements generalized ridge regression, LMM-Lasso implements a two-step lasso approach, and glmnet makes use of 
elastic net for parameter estimation.

BigRR \cite{shen2013novel}  is an R package that implements generalized ridge regression.
In ridge regression, the coefficients of the predictors are shrunk 
to zero by a constant amount.  In generalized ridge regression, the amount of shrinkage is allowed to vary, giving opportunity for 
a more realistic model to be specified. The model used in BigRR can be formulated as a linear mixed model that is 
$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$. 
The parameters are estimated via a 
algorithm that makes clever use of the Cholesky decomposition of the genomic relationship matrix, 
and the equivalence between the above model and animal models \ref{stranden2009animalmodel}. 
One of the challenges in using BigRR for association mapping is that while it has the capacity to estimate the effect sizes of all the snp simultaneously, 
obtaining their statistical significance is difficult. 
Permutation has been suggested as a way of calculating significance 
empirically \cite{shen2013novel} but we instead opted for stability selection.

Lmm-Lasso  \cite{rakitsch2013lasso} is a stand-alone computer program that implements lasso within a linear mixed model 
framework. Unlike in ridge regression, lasso performs parameter estimation and variable selection, simultaneously.   The 
coefficients of "unimportant" predictors are shrunk to exactly zero.
The (full) model is 
$y =  \sum_{l=1}^L m_l \theta_l +   g  + e$ where $g^{(n ; \times 1)}$ is a vector of random polygenic effects such that 
$g \sim N(0, \sigma^2_g K)$, and $K^{(n ; \times \; n)}$ is a realised relationship matrix reconstructed from the snp information. 
It is challenging, computationally,  to apply lasso directly to the full model for parameter estimation. Instead, a two step procedure 
is adopted. In the first step, the above model without any snp effects is fitted. In the second step, 
the unknown variance components in the full model are replaced by their estimates 
that were calculated in the first step. Lasso is then applied to this simpler model making lasso more tractable, computationally. 
As with BigRR,  the statistical significance of the snp effects are not calculated. We use stability selection to obtain the 
significance empirically.

Glmnet  \cite{Friedman2010glmnet}, unlike BigRR and LMM-Lasso, was not purpose built for 
association mapping.  It is a general R package for 
regression analysis. We have included it in our list of applications for comparison with Eagle because the 
regularization method, elastic net, upon which the package is based has been found to be superior to ridge regression and 
lasso (ref needed???).  Elastic net uses a penalty function that is a mixture of 
$\ell_1$-norm (lasso) and $\ell_2$-norm (ridge regression) penalties  \cite{zou2005regularization}. 
This allows elastic net to avoid the limitations of lasso 
while still being able to perform shrinkage and variable selection, simultaneously. 
Glmnet can handle a range of different models. For the analyses performed in this paper, the model is 
$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
We used stability selection to obtain the significance of the coefficients (snp effects).
